Bayesian Sparsification of $\cplx$-valued networks

### 2020-01-15 paper/paper.tex
proof read the text, reworded some phrases

% on dropout
Dropout, \citep{hinton_improving_2012}, prevents overfitting by injecting multiplicative
binary noise into layer's weights, thereby breaking up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
has shown the possibility of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated mutlitask learning, then by sampling uncorrelated weight masks dropout acts
as superposition method, which better utilizes the learning capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex

% segue into bayesian inference
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike bayesian model averaging, which weighs according to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian inference technique. Bayesian Inference as a principled framework of reasoning
about uncertainty, updates prior beliefs in accordance with the likelihood of empirical
observations into posterior belief. Under suitable modelling and prior assumptions, Bayesian
methods can be used towards model sparsification, \citep{kingma_variational_2015,molchanov_variational_2017},
and \citep{kharitonov_variational_2018}.

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015}, $\ell_0$
regularization \citep{louizos_learning_2017} and magnitude pruning on large-scale models.
Their results suggest that although var-dropout may achieve good accuracy-sparsity balance
and outperform pruning on the studied architectures, overall pruning is preferable due to
simplicity and speed. They also conclude that, although dropout and $\ell_0$ yield less
consistent performance and are more complex, they induce non-uniform sparsity throughout
the model, which is essential for better compression \citep{he_amc:_2018}.
\todo{read (He et al., 2018)}
%
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%
Therefore it appears that Bayesian and stochastic sparsification techniques can be useful
for models in $\cplx$ domain. \verify{$\ell_0$ -- stoch?}

@@

The paper is structured as follows. In sec.~\ref{sec:bayesian_dropout} we review Bayesian
dropout techniques, and ins sec.~\ref{sec:c_valued_networks} we provide a brief summary of
the inner working of complex-valued networks as functions of complex argument. The main
contribution of this study is presented in sec.~\ref{sec:dropout_for_c_valued_layers},
where we consider different variational approximations and priors, outline the tricks
and derive expressions for the divergence term. In sec.~\ref{sec:experiments} we evaluate
the sparsification rate and compare the resulting performance of various $\cplx$-networks
proposed in prior work, and discuss the outcomes in sec.~\ref{sec:discussion_and_related_research}.

@@

In general, the core of Bayesian Inference can be summarized as follows: given a prior
distribution $\pi(m)$ on models (hypotheses) $\mathcal{M}$, utilize the empirical evidence
$D$ to update the assumptions by considering the likelihood of the observations under each
$m \in \mathcal{M}$. Models are represented by a parametric family indexed by $\omega \in
\Omega$ and each $m_\omega(\cdot)$ specifies the conditional distribution of the data $
  D = (z_i)_{i=1}^N
$. The posterior distribution $p(\omega \mid D)$, derived using the Bayes rule $
  p(m \mid D) = \tfrac{p(D \mid m) \pi(m)}{p(D)}
$ with $
  p(D) = \mathbb{E}_{\pi(m)} p(D \mid m)
$, provides useful information about yet unobserved data and model parameters, e.g
classification uncertainty, predictive statistics, and parameter relevance.

The posterior itself is intractable, save for the relatively simple cases \citep{citation_needed},
and therefore exact inference is traded for tractability and scalability of Variational
Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach recasts
posterior inference into variational optimization problem: finding $q$ that is close to
$p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \citep{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \citep{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1

Kullback-Leibler divergence of density $q$ from $p$ is
\begin{equation}  \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
and is the natural choice for $\rho$, the main reason being that it satisfies the following
identity for well-behaved densities:
\begin{multline}  \label{eq:kl-div-master}
  % \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
  \log p(D)
    - KL(q \| p(\omega \mid D))
    \\ = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q \| \pi)
    \,.
\end{multline}
This identity yields an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
  \mathcal{L}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \mathbb{E}_{\omega \sim q_{\theta}}
        \log p_{\phi}(D \mid \omega)
      % \mathbb{E}_{\omega \sim q_{\theta}}
        % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % = \mathbb{E}_{\omega \sim q_{\theta}}
  %     \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %   + \mathbb{H}(q_{\theta})
  \,,
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is EM-algorithm. Other
variational objectives are possible, provided $p(\omega \mid D)$ is probed only through
$\log p(D \mid \omega)$ and $\log \pi(\omega)$, \citep{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citep{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased sampling estimators, \citep{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
% (DSVI) titsias-lazaro.tex
\todo{DSVI}
\citep{titsias_doubly_2014}

@@

With special family of posterior approximation, variational inference can be used as
a regularization method and as model sparsification technique. \citet{kingma_variational_2015}
consider Dropout \citep{hinton_improving_2012}, DropConnect, \citep{wan_regularization_2013},
and Gaussian dropout, \citep{srivastava_dropout_2014,wang_fast_2013} through the lens
of Bayesian inference methods and proposes \textit{Variational Dropout}. They argue that
the multiplicative noise introduced by these methods induces a distribution equivalent to
a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $
  \xi_j \sim p_\theta(\xi_j)
$ iid from some $p_\theta(\xi)$.
%
% Binary dropout with rate $p \in (0, 1)$ uses $
%   p_\xi
%     = \mathcal{B}\bigl(
%       \{0, \tfrac1{1-p}\}, 1-p
%     \bigr)
% $, whereas in Gaussian dropout $
%   \xi \sim \mathcal{N}(1, \alpha)
% $ for $\alpha = \tfrac{p}{1-p}$.
% This suggests making $\alpha$ a variational parameter
% and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Thus,
%
Variational Dropout assumes fully factorized Gaussian approximate posterior $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\mu_j, \alpha_j \mu_j^2)
$ and factorized log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. This unravels the divergence term in \eqref{eq:elbo} into a sum of the following
analytically intractable terms:
\begin{equation}  \label{eq:improper-kl-div-real}
  KL(q_\theta \| \pi)
    \propto
      \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\citet{kingma_variational_2015} approximate this expression by a non-linear polynomial
regression for \verify{truncated} $\alpha \in (0, 1)$. \citet{molchanov_variational_2017}
refine the approximation by using sigmoid and soft-plus basis functions.
%
In sec.~\ref{sub:real-chisq-grad} we verify the approximation against Monte-Carlo
estimates of \eqref{eq:improper-kl-div-real} for $\alpha$ varying over a fine logarithmic
grid, and confirm its accuracy, by comparing its derivative to the exact derivative of
\eqref{eq:improper-kl-div-real} with respect to $\log \alpha$.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates

Flexibility in specification of $q_\theta$ enables structured dropout with shared
$\alpha$ across groups of parameters $\omega$. \citet{molchanov_variational_2017} study
Variational Dropout from a model sparsification perspective, optimizing $\alpha$ for each
individual parameter. Another variant of Bayesian dropout, called \textit{Automatic Relevance
Determination}, is developed in \citep{kharitonov_variational_2018}. The key difference is
that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes}
\citep{citation_needed}, fits a prior distribution while performing variational inference.
Maximizing \eqref{eq:elbo} over $\lambda$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta \| \pi)
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

Another contribution of \citep{kingma_variational_2015} is the analysis of the effects of the
\textit{local reparameterization trick} on the variance of the gradient of \eqref{eq:elbo}.
This trick, proposed in \citep{wang_fast_2013} to speed up Gaussian dropout, utilizes the
closure of Gaussians under affine transformations, thereby translating uncertainty from the
global parameter noise to local noise in intermediate outputs within the network.
%
The stochastic output of a linear layer $
  y = W^\top x + b
$ with $
  W \in \mathbb{R}^{n\times m}
$ and independent $
  W_{ij} \sim \mathcal{N}(\mu_{ij}, \alpha_{ij} \mu_{ij}^2)
$ is equivalent in distribution to
\begin{equation}  \label{eq:r-gauss-trick}
    y \sim \prod_i \mathcal{N}\Bigl(
          \sum_j \mu_{ij} x_j + b_i,
          \sum_j \alpha_{ij} \mu_{ij}^2 x_j^2
      \Bigr)
    \,.
\end{equation}
%
\citet{kingma_variational_2015} show, that for the case of fully factorized Gaussian
posterior approximations, the trick decorrelates ELBO estimates within the minibatch,
making the gradient estimator $\nabla_\theta \tilde{\mathcal{L}}$ more statistically
efficient. \citet{molchanov_variational_2017} propose an improvement by introducing
\textit{additive noise parameterization}, which is relevant to model sparsification.
They argue that $(\mu, \alpha)$ parameterization should be abandoned in favour of an
equivalent $(\mu, \sigma^2)$ with the score $\alpha$ calculated using $
  \tfrac{\sigma^2}{\mu^2}
$ when needed. This makes $
  \nabla_\mu \tilde{\mathcal{L}}
$, which is shown in \citep{molchanov_variational_2017} to be independent of the noise,
injected by the trick \eqref{eq:r-gauss-trick}, and further stabilizes gradients in SGVB
\eqref{eq:elbo}.

@@

$\cplx$-Gaussians are closed under affine transformations: for $A \in \cplx^{n \times m}$
and $b \in \cplx^{n}$

@@

% For $\beta = 2$ we add $\gamma$ to make the right-hand side nonegative. (need proof)

@@

The expectation $

@@

$, needed in \eqref{eq:ard-kl-div-real}, by integrating power series \eqref{eq:nonchisq-density}
with substitution $x \to 2u$ equals

@@

Differentiating the power series within $g_m$ yields

@@

using $
  F \colon \mathbb{R} \to \mathbb{R}

@@

$ Dawson's integral (special function).

>>>>

% on dropout
Dropout, \citep{hinton_improving_2012}, prevents overfitting by injecting multiplicative
binary noise into layer's weights, which breaks up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
has shown the possibility of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated mutlitask learning, then by sampling uncorrelated weight masks dropout acts
as superposition method, which better utilizes the learning capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex

% segue into bayesian inference
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike bayesian model averaging, which weighs according to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian Inference technique: a principled framework of reasoning about uncertainty,
whihc updates prior beliefs in accordance with the likelihood of empirical observations
into posterior belief. Under suitable modelling and prior assumptions, Bayesian methods
can be used towards model sparsification, \citep{kingma_variational_2015,molchanov_variational_2017},
and \citep{kharitonov_variational_2018}.

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015}, $\ell_0$
regularization \citep{louizos_learning_2017} and magnitude pruning on large-scale models.
Their results suggest that although var-dropout may achieve good accuracy-sparsity balance
and outperform pruning on the studied architectures, overall pruning is preferable due to
simplicity and speed. They also conclude that, although dropout and $\ell_0$ yield less
consistent performance and are more complex, they induce non-uniform sparsity throughout
the model, which is essential for better compression \citep{he_amc:_2018}.
\todo{read (He et al., 2018)}
%
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%
Therefore it appears that Bayesian and stochastic sparsification techniques can be useful
for models in $\cplx$ domain. \verify{$\ell_0$ -- stoch? yes}

@@

The paper is structured as follows. In sec.~\ref{sec:bayesian_dropout} we review Bayesian
dropout techniques, and in sec.~\ref{sec:c_valued_networks} we provide a brief summary of
the inner working of complex-valued networks as functions of complex argument. The main
contribution of this study is presented in sec.~\ref{sec:dropout_for_c_valued_layers},
where we consider different variational approximations and priors, outline the tricks
and derive expressions for the divergence term. In sec.~\ref{sec:experiments} we evaluate
the sparsification rate and compare the resulting performance of various $\cplx$-networks
proposed in prior work, and discuss the outcomes in sec.~\ref{sec:discussion_and_related_research}.

@@

In general, the core of Bayesian Inference can be summarized as follows: given a prior
distribution $\pi(m)$ on models (hypotheses) $\mathcal{M}$, utilize the empirical evidence
$D$ to update the assumptions by considering the likelihood of the observations under each
$m \in \mathcal{M}$. Models are represented by a parametric family indexed by parameters
$\omega \in \Omega$ and each $m_\omega(\cdot)$ specifies the conditional distribution of
the data $
  D = (z_i)_{i=1}^N
$. The posterior distribution $p(\omega \mid D)$, derived using the Bayes rule $
  p(m \mid D) = \tfrac{p(D \mid m) \pi(m)}{p(D)}
$ with $
  p(D) = \mathbb{E}_{\pi(m)} p(D \mid m)
$, provides useful information about yet unobserved data and model parameters, e.g
classification uncertainty, predictive statistics, and parameter relevance.

The posterior itself is intractable, save for the relatively simple cases \citep{citation_needed},
and therefore exact inference is traded for tractability and scalability offered by Variational
Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach recasts
posterior inference into variational optimization problem: finding $q$ that is close to
$p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \citep{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \citep{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1
%
The natural choice for $\rho$ is Kullback-Leibler divergence of density $q$ from $p$,
\begin{equation}  \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
the main reason being that it satisfies the following identity for well-behaved densities:
% $(q, p, \pi)$
\begin{multline}  \label{eq:kl-div-master}
  % \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
  \log p(D)
    - KL(q \| p(\omega \mid D))
    \\ = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q \| \pi)
    \,.
\end{multline}
This identity yields an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
  \mathcal{L}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \mathbb{E}_{\omega \sim q_{\theta}}
        \log p_{\phi}(D \mid \omega)
      % \mathbb{E}_{\omega \sim q_{\theta}}
        % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % = \mathbb{E}_{\omega \sim q_{\theta}}
  %     \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %   + \mathbb{H}(q_{\theta})
  \,,
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is EM-algorithm. Other
variational objectives are possible, provided $p(\omega \mid D)$ is probed only through
$\log p(D \mid \omega)$ and $\log \pi(\omega)$, \citep{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citep{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which
makes it possible to employ stochastic gradient-based optimization methods to solve
\eqref{eq:variational-progam}. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased sampling estimators, \citep{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
% (DSVI) titsias-lazaro.tex
\todo{DSVI}
\citep{titsias_doubly_2014}

@@

With a special family of posterior approximation, variational inference can be used as
a regularization method and as model sparsification technique. \citet{kingma_variational_2015}
consider Dropout \citep{hinton_improving_2012}, DropConnect \citep{wan_regularization_2013},
and Gaussian dropout \citep{srivastava_dropout_2014,wang_fast_2013} through the lens
of Bayesian inference methods and proposes \textit{Variational Dropout}. They argue that
the multiplicative noise introduced by these methods induces a distribution equivalent to
a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $
  \xi_j \sim p_\theta(\xi_j)
$ iid from some $p_\theta(\xi)$.
%
% Binary dropout with rate $p \in (0, 1)$ uses $
%   p_\xi
%     = \mathcal{B}\bigl(
%       \{0, \tfrac1{1-p}\}, 1-p
%     \bigr)
% $, whereas in Gaussian dropout $
%   \xi \sim \mathcal{N}(1, \alpha)
% $ for $\alpha = \tfrac{p}{1-p}$.
% This suggests making $\alpha$ a variational parameter
% and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Thus,
%
Variational Dropout assumes fully factorized Gaussian approximate posterior $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\mu_j, \alpha_j \mu_j^2)
$ and factorized log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. The divergence term in \eqref{eq:elbo} unravels into a sum of the following
analytically intractable terms:
\begin{equation}  \label{eq:improper-kl-div-real}
  KL(q_\theta \| \pi)
    \propto
      \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\citet{kingma_variational_2015} approximate this expression by a non-linear polynomial
regression for truncated $\alpha \in (0, 1)$. \citet{molchanov_variational_2017}
refine the approximation by using a nonlinear regression based on sigmoid and soft-plus.
% see p.6 sec 3.4 of kingma_variational_2015
%
In sec.~\ref{sub:real-chisq-grad} we verify the approximation against Monte-Carlo
estimates of \eqref{eq:improper-kl-div-real} for $\alpha$ varying over a fine logarithmic
grid, and confirm its accuracy, by comparing its derivative to the exact derivative of
\eqref{eq:improper-kl-div-real} with respect to $\log \alpha$.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates

Being the ratio of the squared mean to the effective variance, $\alpha_j$ regulates the
relevance of the parameter $\omega_j$ it is associated to. Thus \citet{molchanov_variational_2017}
focus on the model sparsification capabilities of Variational Dropout, optimizing $\alpha$
for each individual parameter. Flexibility in specification of $q_\theta$ enables structured
dropout with shared $\alpha$ across groups of parameters $\omega$. Another variant of Bayesian
dropout, called \textit{Automatic Relevance Determination}, is developed in \citep{kharitonov_variational_2018}.
The key difference is that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper
Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes}
\citep{citation_needed}, fits a prior distribution while performing variational inference.
Maximizing \eqref{eq:elbo} over $\lambda$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta \| \pi)
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

Another contribution of \citep{kingma_variational_2015} is the analysis of the effects of the
\textit{local reparameterization trick} on the variance of the gradient of \eqref{eq:elbo}.
This trick, proposed in \citep{wang_fast_2013} to speed up Gaussian dropout, utilizes the
closure of Gaussians under affine transformations, thereby translating uncertainty from the
global parameter noise to local noise in intermediate outputs within the network.
%
The stochastic output of a linear layer $
  y = W^\top x + b
$ with $
  W \in \mathbb{R}^{n\times m}
$ and independent $
  W_{ij} \sim \mathcal{N}(\mu_{ij}, \alpha_{ij} \mu_{ij}^2)
$ is equivalent in distribution to
\begin{equation}  \label{eq:r-gauss-trick}
    y \sim \prod_i \mathcal{N}\Bigl(
          \sum_j \mu_{ij} x_j + b_i,
          \sum_j \alpha_{ij} \mu_{ij}^2 x_j^2
      \Bigr)
    \,.
\end{equation}
%
\citet{kingma_variational_2015} show, that for the case of fully factorized Gaussian
posterior approximations, the trick decorrelates ELBO estimates within the minibatch,
making the gradient estimator $\nabla_\theta \tilde{\mathcal{L}}$ more statistically
efficient. \citet{molchanov_variational_2017} propose an improvement by introducing
\textit{additive noise parameterization}. They argue in favour of abandoning the
$(\mu, \alpha)$ parameterization in favour of an equivalent $(\mu, \sigma^2)$, since
the latter renders $
  \nabla_\mu \tilde{\mathcal{L}}
$ independent from the local output noise, injected by the trick \eqref{eq:r-gauss-trick},
and further stabilizes gradients in SGVB \eqref{eq:elbo}. The relevance $\alpha$ is
calculated using $
  \tfrac{\sigma^2}{\mu^2}
$ when needed.

@@

$\cplx$-Gaussian random variables are closed under affine transformations: for $
  A \in \cplx^{n \times m}
$ and $b \in \cplx^{n}$

@@

% For $\beta = 2$ we add $\gamma$ to make the right-hand side nonegative. (proof needed)

@@

By integrating the power series \eqref{eq:nonchisq-density} with substitution $x \to 2u$,
the expectation $

@@

$, needed in \eqref{eq:improper-kl-div-real}, equals

@@

Formally differentiating the convergent power series within $g_m$ yields

@@

using Dawson's integral $
  F
  % \colon \mathbb{R} \to \mathbb{R}

@@
